---
title: Food Safety Analysis of Subway and it's Two Largest Competitors (Jimmy Johns
  and Quiznos) In the state of Maryland
author: "Ethan McGregor"
output:
  html_document: default
  pdf_document: default
---


```{r setup, echo=FALSE, message=FALSE, warning=FALSE}

library('plotly')
library('dplyr')
library('stringr')
library('leaflet')
require('knitr')

# import data from data.gov
data <- data.frame(read.csv("Food_inspection.csv"))

```

```{r get_each_company_data, echo=FALSE, message=FALSE, warning=FALSE}
# Filtering our data down to the three compnaies we are looking at
# Data (Name)is not consistant so we have to partial filter by names

jimmy <- (data %>% filter(str_detect(Name,"JIMMY")))

subway <- (data %>% filter(str_detect(Name,"SUBWAY")))
                          
quiznos <- (data %>% filter(str_detect(Name,"QUIZ")))

```



Subway is the largest sandwich company in Maryland with `r length(unique(subway$Location))` locations with Jimmy Johns trailing behind with `r length(unique(jimmy$Location))` and Quiznos with only `r length(unique(quiznos$Location))`

```{r Subway_Map, echo=FALSE, message=FALSE, warning=FALSE}

# changes inspection results to a binary 0 or 1 with 0 being no critical violations
subway$binary_inspection_results <- ifelse(subway$Inspection.Results == "No Critical Violations Noted",0,1)
quiznos$binary_inspection_results <- ifelse(quiznos$Inspection.Results == "No Critical Violations Noted",0,1)
jimmy$binary_inspection_results <- ifelse(jimmy$Inspection.Results == "No Critical Violations Noted",0,1)

# new var to use when hover mouse over map shows sum of data
subway$hover <- with(subway, paste("City: ", City, ", Inspection Results: ", Inspection.Results, "Lat: ", Latitude, "long", Longitude))

# creates interactive map
pal <- colorFactor(c("Blue","Red"),
  domain = 0:1)

m <- leaflet(data = subway) %>%
  addTiles() %>% 
  addCircleMarkers(~Longitude, ~Latitude, color = ~pal(binary_inspection_results),

label = ~hover)


```

#Subway locations and Inspection status
+ RED = Violation Corrected
+ BLUE = No critical Violations
+ Darker the color = more inspections
+ Hover for more information
`r m`


##Health Code Violation Index Score
###(By Zipcode)
####Higher the number (0-1) means more of their inspections resulted in a violation in that zipcode
```{r subwaybarChart, echo=FALSE, message=FALSE, warning=FALSE}
# original bar chart for subway but moved to new grouped chart
#p <- plot_ly(
#  x = factor(subwayScore$Zip),
 # y = subwayScore$SScore,
 # name = "Rating by Zipcode",
 # type = "bar"
#)
```


```{r groupedBarChart , echo=FALSE, message=FALSE, warning=FALSE}

# adds a score column which is the mean of the binary violations, if a higher number (closer to 1) is found more of their
# inspections resulted in a violation
jimmyScore <- summarise(group_by(jimmy, Zip), JScore = mean(binary_inspection_results, na.rm = TRUE))
quiznosScore <- summarise(group_by(quiznos, Zip), QScore = mean(binary_inspection_results, na.rm = TRUE))
subwayScore <- summarise(group_by(subway, Zip), SScore = mean(binary_inspection_results, na.rm = TRUE))

# joins scores into new dataframe for use in the chart
twoCombinedScores <-right_join(jimmyScore, subwayScore)
allCombinedScores <- left_join(twoCombinedScores,quiznosScore)

groupedGraph <- plot_ly(allCombinedScores, x = factor(allCombinedScores$Zip), y = ~JScore, type = 'bar', name = 'Jimmy Johns') %>%
  add_trace(y = ~SScore, name = 'Subway') %>% 
  add_trace(y= ~QScore, name = 'Quiznos') %>%
  layout(yaxis = list(title = 'Count'), barmode = 'group')%>% 
    layout(xaxis = list(tickangle = 45))


```
`r groupedGraph`

  While Subway has many more locations than Quiznos or Jimmy Johns we can see that their scores are not abnormal from their competitors. While we can see there are a few areas specifically 20853 and that have a high amount of problems, we can use that knowledge to target food saftey training in that area.
  
```{r pieChart, echo=FALSE, message=FALSE, warning=FALSE}

inspectionTypeCounts <- count(group_by(subway, Inspection.Type)) %>% filter(Inspection.Type %in% c("Complaint Invest w/Comprehensive","Comprehensive","Food Complaint","Monitoring","Monitoring w/Complaint Investigation"))

inspectionReasonPieChart <- plot_ly(inspectionTypeCounts, labels = ~Inspection.Type, values = ~n, type = 'pie') %>%
  layout(title = 'Reasons for Inspections (Subway)',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))



```

`r inspectionReasonPieChart`

The vast majority of Inspections at Subway have been routine, either monitoring or Comprehensive. Less then 3 percent of all inspections have been because of a complaint. 


